Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/eval.tex
x

To evaluate how much programmer effort is required to build an
application that uses Nail, we implemented three applications---a
DNS server, a replacement for the \cc{host} DNS client tool, and an
\cc{unzip} program---based on the above grammars, and compared code
size with comparable applications that process data manually, using
SlocCount~\cite{sloccount}.  We also compare the code size of our
DNS server to a DNS server written using the Hammer parsing framework,
although it does not fully support DNS (e.g., it lacks label compression,
among other things).  Figure~\ref{fig:effort} summarizes the results.

\XXX[nz][should we kill \cc{host}, since our replacement is nowhere
near as full-featured as the real \cc{host} or \cc{dig}?]

\begin{figure}[tb]
\centering
\begin{tabular}{llr@{~}l}
\toprule
\textbf{System}
  & \textbf{LoC w/ Nail}
  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
\midrule
DNS server
  & 295
  & 683
  & (Hammer parser) \\

@@

DNS client
  & 210
  & 4,615
  & (host) \\

@@

Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full},
consists of a 48 line Nail grammar and 64 lines of C
implementing DNS label compression, whose signatures are shown in
Figure~\ref{fig:dns-xform}.  The grammar describes both the structure
of DNS packets (36 lines) and of a simple zone-file format (16 lines)
supporting A, NS, MX and CNAME records.

With this grammar, we built a simple authoritative DNS server, which parses a zone file, listens to incoming DNS
requests, parses them and generates appropriate responses, is implemented in 183 lines of C. The
same grammar is used, together with 98 lines of C, to implement a minimal clone of the
\cc{host} command-line tool. Most of this code consists of a custom hashtable and system
interface code, such as listening to sockets and reading the zonefile.

The Hammer parser framework\cite{hammer-parser} ships with a toy DNS server that responds to
any valid DNS query with a CNAME record to the domain ``spargelze.it''. 
The server consists of 683 lines of C, mostly custom validators, semantic actions,
and data structure definitions, with only 52 lines of code defining the
grammar with Hammer's combinators.

It is hard to compare the programming effort required to implement
our DNS server to that of a real world DNS server, since we implement less functionality.
 In particular, we do not send the additional glue records real-world DNS servers send and
we did not implement any configuration options.

However, the closest in functionality and intent is Dan Bernstein's
djbdns,\footnote{\url{http://cr.yp.to/djbdns.html}} which aims to be a minimalist, highly secure,
authoritative-only DNS server. The latest release of djbdns, including various support tools, is
about 10,000 lines of C as measured by \cc{sloccount}. DJ

>>>>

To evaluate how much programmer effort is required to build an
application that uses Nail, we implemented two applications---a
DNS server, and an
\cc{unzip} program---based on the above grammars, and compared code
size with comparable applications that process data manually, using
\cc{sloccount}~\cite{sloccount}.  We also compare the code size of our
DNS server to a DNS server written using the Hammer parsing framework,
although it does not fully support DNS (e.g., it lacks label compression,
among other things).  Figure~\ref{fig:effort} summarizes the results.

\begin{figure}[tb]
\centering
\begin{tabular}{@{~}lrr@{~}l@{~}}
\toprule
\textbf{Application}
  & \textbf{LoC w/ Nail}
  & \multicolumn{2}{c}{\textbf{LoC w/o Nail}} \\
\midrule
DNS server
  & 295
  & 683
  & (Hammer parser) \\

@@

@@

Our DNS server parses a zone file, listens to incoming DNS requests,
parses them, and generates appropriate responses.  The DNS server is
implemented in 183 lines of C, together with 48 lines of Nail grammar
and 64 lines of C code implementing stream transforms for DNS label
compression.  In comparison, Hammer~\cite{hammer-parser} ships with a toy
DNS server that responds to any valid DNS query with a CNAME record to the
domain ``spargelze.it''.  Their server consists of 683 lines of C, mostly
custom validators, semantic actions, and data structure definitions, with
only 52 lines of code defining the grammar with Hammer's combinators.
Their DNS server does not implement label compression, zone files, etc.
From this, we conclude that Nail leads to much more compact code for
dealing with DNS packet formats.

It is hard to compare the programming effort required to implement
our DNS server to that of a real world DNS server, since we implement
less functionality: we do not send the additional glue records and
we did not implement any configuration options.  However, the closest
in functionality and intent is Dan Bernstein's djbdns~\cite{djbdns},
which aims to be a minimalist, highly secure, authoritative-only DNS
server. The latest release of djbdns, including various support tools,
is about 10,000 lines of C\@.  \XXX[nz][so what does this mean?]
